Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-21 paper.qmd
Updated Figure 2

@@

projection <- cbind(
    c(0.24288,-0.03628,-0.03302,0.12182,-0.04150,0.02344,0.05910),
    c(0.05019,0.07165,-0.20131,-0.16200,-0.05343,0.06854,-0.00551))

@@

@@

@@

>>>>

```{r}
#| label: scale-highd-data
# Center the data by subtracting the mean of each column
center_data <- function(data) {
  apply(data, 2, function(col) col - mean(col))
}

# Function to scale data manually
scale_data_manual <- function(data) {
  # Step 1: Center the data (mean 0)
  data_centered <- center_data(data)

  # Step 2: Calculate the range of each dimension
  ranges <- apply(data_centered, 2, function(x) max(x) - min(x))

  # Step 3: Find the dimension with the largest range
  max_range <- max(ranges)
  scaled_data <- data_centered

  # Step 4: Scale each dimension
  for (i in seq_along(ranges)) {
    scaling_factor <- ranges[i] / max_range
    scaled_data[, i] <- data_centered[, i] / ranges[i] * scaling_factor
  }

  # Step 5: Rescale the dimension with the largest range to [-1, 1]
  largest_dim_index <- which.max(ranges)
  scaled_data[, largest_dim_index] <- scaled_data[, largest_dim_index] * max_range

  return(scaled_data)
}
```

@@

# Apply the scaling function to the s_curve_noise data
scaled_s_curve_noise <- scale_data_manual(
  training_data_scurve |> select(-ID)) |>
  as_tibble()

# Apply the scaling function to the s_curve_noise data
scaled_s_curve_true <- scale_data_manual(
  true_model_df |> select(-ID)) |>
  as_tibble()

# projection <- cbind(
#     c(0.24288,-0.03628,-0.03302,0.12182,-0.04150,0.02344,0.05910),
#     c(0.05019,0.07165,-0.20131,-0.16200,-0.05343,0.06854,-0.00551))

projection <- cbind(
  c(0.5898,-0.4620,-0.1362,0.2047,-0.3680,0.1032,-0.4818),
  c(0.1756,0.0257,0.3683,0.2400,-0.5518,-0.4573,0.5115))

projected <- as.matrix(scaled_s_curve_noise) %*% projection

@@

projected_true <- as.matrix(scaled_s_curve_true) %*% projection

@@

limits <- 1
rng <- range(projected)
projected <- projected/max(abs(rng))
colnames(projected) <- c("P1", "P2")
projected <- data.frame(projected)
obs_labels <- as.character(1:nrow(scaled_s_curve_noise))

@@

axis_labels <- colnames(scaled_s_curve_noise)
rownames(axes) <- axis_labels
